Clustering in a hyperbolic model of complex networks.

### 2019-02-22 concentration.tex
revised section 8.1.

In this section we establish a concentration result for the local clustering function $c^\ast(k; \Gbox)$ in the finite model $\Gbox$. Similar to the previous section we will focus on typical points $p = (0,y)$ with $y \in \Kcal_{C}(k_n)$.

@@

Recall the adjusted triangle count function \eqref{eq:def_tilde_triangle_indicator}
\[
	\widetilde{T}_{\Pcal,n}(p_0,p_1,p_2) = \ind{p_1 \in \BallPon{p}}\ind{p_2 \in \BallPon{p}}\ind{p_2 \in \BallPo{p_1} \cap \Rcal}.
\]
and the concentration set
\[
	\Kcal_C(k_n,n) = \left\{p \in \Rcal : \frac{k_n - C \kappa_n}{\xi} \vee 0 \le e^{\frac{y}{2}}
		\le \frac{k_n + C \kappa_n}{\xi} \right\},
\]
with
\[
	\kappa_n = \sqrt{k_n \log(k_n)}.
\]
We will define the corresponding triangle degree function
\begin{equation}\label{eq:def_degree_triangle_count_in_K}
	\widetilde{T}_{\Pcal,n}(k, C) = \sum_{p \in \Pcal_n \cap \Kcal_{C,n}(k_n)} \ind{\text{deg}_{\text{box}}(p) = k} \widetilde{T}_{\Pcal,n}(p),
\end{equation}
where 
\[
	\widetilde{T}_{\Pcal,n}(p) := \sum_{(p_1, p_2) \in \Pcal \setminus p}^{\ne} \widetilde{T}_{\Pcal,n}(p,p_1,p_2),
\]
where the sum is over all distinct pairs $(p_1, p_2) \in \Pcal \setminus p$.

Next we note that the second result from Corollary~\ref{cor:adjusted_triangle_counting_P_n} states that
\begin{equation}\label{eq:asymptotic_equivalence_contraint_triangle_count}
	\Exp{\widetilde{T}_{\Pcal,n}(k_n,C)} = (1+\smallO{1})\Exp{T_{\Pcal}(k_n)},
\end{equation}
so that we can conclude that the main contribution of triangles of degree $k_n$ is given by $\widetilde{T}_{\Pcal,n}(k_n, C)$. Therefore, in order to prove Proposition \ref{prop:concentration_local_clustering_P_n} it suffices to show that $\widetilde{T}_{\Pcal,n}(k_n,C)$ is sufficiently concentrated around its mean. This is done in the following proposition.

%\subsection{The proof of Proposition \ref{prop:concentration_local_clustering_P_n}}
%
%We start with the concentration result for $\widetilde{T}_{\Pcal,n}(k_n)$.

\begin{proposition}[Concentration $\widetilde{T}_{\Pcal,n}(k_n,C)$]\label{prop:concentration_tilde_T_P_n}
Let $\alpha > \frac{1}{2}$, $\nu > 0$ and let $(k_n)_{n \ge 1}$ be any positive sequence satisfying $k_n = \smallO{n^{\frac{1}{2\alpha+1}}}$. Then for any $C > 0$, as $n \to \infty$,
\[
	\Exp{\widetilde{T}_{\Pcal,n}(k_n,C)^2} = \left(1 + \smallO{1}\right)\Exp{\widetilde{T}_{\Pcal,n}(k_n, C)}^2.
\]
\end{proposition}

We postpone the proof of this proposition postpone till Section \ref{ssec:concentration_tilde_T} and first use it to prove Proposition \ref{prop:concentration_local_clustering_P_n}.

%\begin{lemma}\label{lem:L1_convergence_c_Pcal_n}
%Let $\nu > 0$, $\alpha > \frac{1}{2}$ and $k_n = o\left(n^{\frac{1}{2\alpha + 1}}\right)$. Then, as $n \to \infty$,
%\[
%	\Exp{\left|c^\ast(k_n; \Gbox) - c_{\Pcal, n}(k_n)\right|} = \smallO{\Exp{c_{\Pcal,n}(k_n)^\ast}}.
%\]
%\end{lemma}
%
%\begin{proof}
%Let $0 < \delta < 1$ and define the following two events
%\begin{align*}
%	A_n &= \left\{\left|N_{\text{box}}(k_n) - \Exp{N_{\text{box}}(k_n)}\right| \le \Exp{N_{\text{box}}(k_n)}^{\frac{1 + \delta}{2}}\right\}\\
%	B_n &= \left\{\left|N_{\text{box}}(k_n) - \Exp{N_{\text{box}}(k_n)}\right| \le n\right\},
%\end{align*}
%so that
%\[
%	1 = \ind{A_n} + \ind{A_n^c}\ind{B_n} + \ind{B_n^c}.
%\]
%Next we note that on the event $A_n$
%\[
%	\left|\frac{\Exp{N_{\text{box}}(k_n)}}{N_{\text{box}}(k_n)} - 1\right| 
%	\le \frac{\Exp{N_{\text{box}}(k_n)}^{\frac{1 + \delta}{2}}}{\Exp{N_{\text{box}}(k_n)}+\Exp{N_{\text{box}}(k_n)}^{\frac{1 + \delta}{2}}}
%	\le \Exp{N_{\text{box}}(k_n)}^{-\frac{1 - \delta}{2}},
%\]
%and on the event $B_n$
%\[
%	\left|\frac{\Exp{N_{\text{box}}(k_n)}}{N_{\text{box}}(k_n)} - 1\right| 
%	\le \frac{n}{\Exp{N_{\text{box}}(k_n)} + n} \le 1.
%\]
%Therefore we have
%\begin{align*}
%	\Exp{\left|c^\ast(k_n; \Gbox) - c_{\Pcal, n}(k_n)\right|}
%	&= \Exp{c^\ast(k_n; \Gbox)\left|\frac{\Exp{N_{\text{box}}(k_n)}}{N_{\text{box}}(k_n)} - 1\right|}\\
%	&\le \Exp{c^\ast(k_n; \Gbox)\ind{A_n}}\Exp{N_{\text{box}}(k_n)}^{-\frac{1 - \delta}{2}} \\
%	&\hspace{10pt}+ \Exp{c^\ast(k_n; \Gbox)\ind{A_n^c}\ind{B_n}} \\
%	&\hspace{10pt}+ \Exp{c^\ast(k_n; \Gbox)\left|\frac{\Exp{N_{\text{box}}(k_n)}}{N_{\text{box}}(k_n)} - 1\right|\ind{B_n^c}}
%\end{align*}
%Since $\Exp{N_{\text{box}}(k_n)} \to \infty$, the first term is clearly $\smallO{\Exp{c^\ast(k_n; \Gbox)}}$. For the third term we have
%\begin{align*}
%	\Exp{c^\ast(k_n; \Gbox)\left|\frac{\Exp{N_{\text{box}}(k_n)}}{N_{\text{box}}(k_n)} - 1\right|\ind{B_n^c}} 
%	&= \bigO{\Prob{B_n^c}}
%		= \bigO{\Prob{\left|N_{\text{box}}(k_n) - \Exp{N_{\text{box}}(k_n)}\right| > n}}\\
%	&= \bigO{e^{-\frac{n^2}{\Exp{N_{k_n}} + n}}} = \bigO{e^{-n}} = \smallO{\Exp{c^\ast(k_n; \Gbox)}}.
%\end{align*}
%Hence we are left to show that
%\begin{equation}\label{eq:L1_convergence_second_term}
%	\Exp{c^\ast(k_n; \Gbox)\ind{A_n^c}\ind{B_n}} = \smallO{\Exp{c^\ast(k_n; \Gbox)}}.
%\end{equation}
%By writing
%\begin{align*}
%	c^\ast(k_n; \Gbox) = \frac{\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}}
%	+ \frac{T_{\Pcal,n}(k_n,\varepsilon)- \widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}}
%\end{align*}
%and using Lemma~\ref{lem:clustering_error_T_term} we get
%\begin{align*}
%	\Exp{c^\ast(k_n; \Gbox)\ind{A_n^c}\ind{B_n}}
%	&\le \Exp{\frac{(1+\smallO{1})\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}}
%		\ind{A_n^c}} + \smallO{\Exp{c^\ast(k_n; \Gbox)}}.
%\end{align*}
%For the last step we use H\"{o}lder's inequality and Proposition \ref{prop:concentration_tilde_T_P_n} to get
%\begin{align*}
%	\Exp{\frac{\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}}\ind{A_n^c}}
%	&\le \Exp{\left(\frac{\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}
%		{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}}\right)^2}^{\frac{1}{2}} \Prob{A_n^c}^{\frac{1}{2}}\\
%	&= \left(1 + \smallO{1}\right) 	
%		\frac{\Exp{\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}}{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}}
%		\Prob{A_n^c}^{\frac{1}{2}}\\
%	&= \bigO{\Exp{c^\ast(k_n; \Gbox)}}\Prob{A_n^c}^{\frac{1}{2}} = \smallO{\Exp{c^\ast(k_n; \Gbox)}},
%\end{align*}
%since $\Prob{A_n} = o(1)$. This establishes \eqref{eq:L1_convergence_second_term} and hence we conclude that
%\[
%	\Exp{\left|c^\ast(k_n; \Gbox) - c_{\Pcal, n}(k_n)\right|} = \smallO{\Exp{c_{\Pcal,n}(k_n)^\ast}}.
%\]
%\end{proof}

\begin{proof}[Proof of Proposition \ref{prop:concentration_local_clustering_P_n}]
Again, we write
\begin{align*}
	c^\ast(k_n; \Gbox) = \frac{\widetilde{T}_{\Pcal,n}(k_n,C)}{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}}
	+ \frac{\left(T_{\Pcal,n}(k_n)-\widetilde{T}_{\Pcal,n}(k_n,C)\right)}{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}},
\end{align*}
so that by~\eqref{eq:asymptotic_equivalence_contraint_triangle_count},
\begin{equation}\label{eq:concentration_local_clustering_P_n_main}
	\Exp{c^\ast(k_n; \Gbox)} = \frac{\Exp{\widetilde{T}_{\Pcal,n}(k_n,C)}}{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}}
	+ \smallO{\Exp{c^\ast(k_n; \Gbox)}}.
\end{equation}
%Therefore, by Lemma \ref{lem:L1_convergence_c_Pcal_n},
%\begin{align*}
%	\Exp{\left|c_{\Pcal,n}(k_n) - \Exp{c^\ast(k_n; \Gbox)}\right|}
%	&\le \Exp{\left|c^\ast(k_n; \Gbox) - \Exp{c^\ast(k_n; \Gbox)}\right|}
%		+ \Exp{\left|c_{\Pcal,n}(k_n) - c^\ast(k_n; \Gbox)\right|}\\
%	&= \frac{\Exp{\left|\widetilde{T}_{\Pcal,n}(k_n,\varepsilon) - \Exp{\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}\right|}}
%		{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}} + \smallO{\Exp{c^\ast(k_n; \Gbox)}}.
%		\numberthis \label{eq:concentration_local_clustering_P_n_term1}
%\end{align*}
Next, we use Proposition \ref{prop:concentration_tilde_T_P_n} to obtain
\begin{align*}
	\Exp{\left|\widetilde{T}_{\Pcal,n}(k_n,C) - \Exp{\widetilde{T}_{\Pcal,n}(k_n,C)}\right|}
	&\le \left(\Exp{\widetilde{T}_{\Pcal,n}(k_n,C)^2} 
		- \Exp{\widetilde{T}_{\Pcal,n}(k_n,C)}^2\right)^{\frac{1}{2}}\\
	&= \smallO{\Exp{\widetilde{T}_{\Pcal,n}(k_n,C)}}.
\end{align*}
This implies
\begin{align*}
	\frac{\Exp{\left|\widetilde{T}_{\Pcal,n}(k_n,C) - \Exp{\widetilde{T}_{\Pcal,n}(k_n,C)}\right|}}
		{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}}
	&= \smallO{\Exp{c^\ast(k_n; \Gbox)}},
\end{align*}
which together with~\eqref{eq:concentration_local_clustering_P_n_main} yields the required result.
\end{proof}

@@

Recall (see \eqref{eq:def_degree_triangle_count_in_K}) that for any $C > 0$
\[
	\widetilde{T}_{\Pcal,n}(k,C) = \sum_{p \in \Pcal_n \cap \Kcal_{C,n}(k_n)} \ind{\text{deg}_{\text{box}}(p) = k} \widetilde{T}_{\Pcal,n}(p)
\]
Then we have
\[
	\widetilde{T}_{\Pcal,n}(k,C)^2 = \hspace{-5pt} \sum_{p, p^\prime \in \Pcal_n \cap \Kcal_C(k)}
		\hspace{-3pt} \ind{D_{\Pcal,n}(p), \, D_{\Pcal,n}(p^\prime) = k} 
		\sum_{(p_1, p_2), (p_1^\prime, p_2^\prime) \in \Pcal_n \setminus (0,y)}^{\ne} \hspace{-3pt}
		\widetilde{T}_{\Pcal}(p,p_1,p_2) \widetilde{T}_{\Pcal}(p^\prime, p_1^\prime, p_2^\prime),
\]
This expression can be written as the sums of several terms, depending on how $\{p, p_1, p_2\}$ and $\{p^\prime, p_1^\prime, p_2^\prime\}$ intersect. To this end we define, for $a \in \{0,1\}$ and $b \in \{0,1,2\}$,
\[
	I_{a,b} = \hspace{-3pt} \sum_{p, p^\prime \in \Pcal_n \cap \Kcal_C(k) \atop |\{p\} \cap \{p^\prime\}| = a}
	\hspace{-5pt} \ind{D_{\Pcal,n}(p), \, D_{\Pcal,n}(p^\prime) = k} J_b(p,p^\prime),
\]
where
\[
	J_b(p,p^\prime) = \hspace{-10pt} \sum_{p_1, p_2, p_1^\prime, p_2^\prime \in \Pcal_n 
		\atop |\{p_1, p_2\} \cap \{p_1^\prime, p_2^\prime\}| = b}^{\ne}
		\hspace{-5pt} T_{\Pcal,n}(p,p_1,p_2) T_{\Pcal,n}(p^\prime, p_1^\prime, p_2^\prime),
\]
with the summation being over all two distinct pairs $(p_1, p_2)$ and $(p_1^\prime, p_2^\prime)$. Then we have
\[
	\widetilde{T}_{\Pcal,n}(k, C)^2 = \sum_{a = 0}^1 \sum_{b = 0}^2 I_{a,b}.
\]

To prove Proposition \ref{prop:concentration_tilde_T_P_n} we will deal with each of the $I_{a,b}$ separately, showing that 
\begin{equation}\label{eq:variance_T_I_00}
	\Exp{I_{0,0}} = (1+\smallO{1})\Exp{\widetilde{T}_{\Pcal,n}(k_n)}^2
\end{equation}
and for all other combinations
\begin{equation}\label{eq:variance_T_I_ab}
	\Exp{I_{a,b}} = \smallO{\Exp{\widetilde{T}_{\Pcal,n}(k_n)}^2}.
\end{equation}
Note that $J_{b}(p,p^\prime) \le J_{0}(p,p^\prime)$ and, since $I_{1,2} = \widetilde{T}_{\Pcal,n}(k_n,C)$, \eqref{eq:variance_T_I_ab} holds for $I_{1,2}$.

@@

Throughout this proof we set $i = |\{p^\prime, p_1, p_2, p_1^\prime, p_2^\prime\} \cap \BallPon{p}|$, $j = |\{p^\prime\} \cap \BallPon{p}|$ and define $i^\prime, j^\prime$ in a similar way by interchanging the primed and non-primed variables. In addition, we write $D_{\Pcal,n}(p,p^\prime,k,\ell)$ to denote the indicator that $|\BallPon{p} \cap (\Pcal_n \setminus \{p, p^\prime, p_1, p_2, p_1^\prime, p_2^\prime\})| = k$ and $|\BallPon{p^\prime} \cap (\Pcal_n \setminus \{p, p^\prime, p_1, p_2, p_1^\prime, p_2^\prime\})| = \ell$. Then, 
\begin{align*}
	&\Exp{\ind{D_{\Pcal,n}(p) = k_n, D_{\Pcal,n}(p^\prime) = k_n} J_b(p,p^\prime)}\\
	&= \Exp{\sum_{p_1, p_2, p_1^\prime, p_2^\prime \in \Pcal_n 
		\atop |\{p_1, p_2\} \cap \{p_1^\prime, p_2^\prime\}| = b}^{\ne}
			\hspace{-7pt} D_{\Pcal,n}(p,p^\prime, k_n-i,k_n- i^\prime) \,
			\widetilde{T}_{\Pcal,n}(p,p_1,p_2) \widetilde{T}_{\Pcal,n}(p^\prime, p_1^\prime, p_2^\prime)},
\end{align*}
where the sum is over all distinct pairs $(p_1, p_2)$ and $(p_1^\prime, p_2^\prime)$. We also that 
\[
	\Exp{T_\Pcal(k_n)} = \bigT{n k_n^{-(2\alpha - 1)} s_\alpha(k_n)}.
\]

@@

%	&\le \int_{\mathcal{E}_{C}(k_n)^c} \rho_{n}(y,k_n) \Exp{\widetilde{T}_{\Pcal,n}(p)}
%		\Exp{\widetilde{T}_{\Pcal,n}(p^\prime)} 

@@

%	&\le \bigO{\log(n)^2} \left(\int_{\Kcal_{C}(k_n)} \rho_{n}(y,k_n) \Exp{\widetilde{T}_{\Pcal,n}(p)} 

@@

%	\mathcal{J}_1^\ast = \int_{\Rcal^2} \ind{0\le y_1, y_2 \le h(p,p^\prime)} \widetilde{T}_{\Pcal,n}(p,p_1,p_2) 

@@

%	\mathcal{J}_1^\ast = (1+\smallO{1})\Exp{\widetilde{T}_{\Pcal,n}(p)} = (1 + \smallO{1})\binom{k_n}{2} \Delta_\Pcal(y),

@@

By Lemma~\ref{lem:joint_degree_distribution_shift} 
\[
	\rho_(p,p^\prime, k_n-i,k_n-i^\prime) = (1+o(1))\rho(p,p^\prime, k_n-j,k_n-j^\prime) 
	= (1+o(1))\rho_{n}(p, p^\prime, k_n, k_n)
\]
which now no longer depends on the other four points $p_1, p_2, p_1^\prime, p_2^\prime$. Hence, using the Campbell-Mecke formula, we get
\[
	\Exp{I_{0,0}^\ast} = (1 + \smallO{1})\int_{\mathcal{E}_{\varepsilon}(k_n)}\rho_{n}(p,p^\prime,k_n, k_n)
		\Exp{\widetilde{T}_{\Pcal,n}(p)}\Exp{\widetilde{T}_{\Pcal,n}(p^\prime)} f(x,y)
		f(x^\prime,y^\prime) \, dx^\prime \, dx \, dy^\prime \, dy,
\]

>>>>

In this section we establish a concentration result for the local clustering function $c^\ast(k; \Gbox)$ in the finite box model $\Gbox$. Similar to the previous section we will focus on typical points $p = (0,y)$ with $y \in \Kcal_{C}(k_n)$.

@@

Recall the adjusted triangle count function
\[
	\widetilde{T}_{\text{box}}(p_0) = \sum_{(p_1, p_2) \in \Pcal \setminus p_0}^{\ne} 
		\widetilde{T}_{\text{box}}(p_0,p_1,p_2).
\]
where
\[
	\widetilde{T}_{\text{box}}(p_0,p_1,p_2) = \ind{p_1 \in \BallPon{p}}\ind{p_2 \in \BallPon{p}}\ind{p_2 \in \BallPo{p_1} \cap \Rcal},
\]
and recall the definition of $\Kcal_{C,n}(k_n)$
\[
	\Kcal_{C,n}(k_n) = \left\{p \in \Rcal : \frac{k_n - C \sqrt{k_n \log(k_n)}}{\xi} \vee 1 \le e^{\frac{y}{2}}
		\le \frac{k_n + C \sqrt{k_n \log(k_n)}}{\xi} \right\}.
\]

Slightly abusing notation, we will define the corresponding triangle degree function
\begin{equation}\label{eq:def_degree_triangle_count_in_K}
	\widetilde{T}_{\text{box}}(k, C) = \sum_{p \in \Pcal_n \cap \Kcal_{C,n}(k_n)} \ind{\text{deg}_{\text{box}}(p) = k} \widetilde{T}_{\text{box}}(p).
\end{equation}
The idea is that the main contribution of triangles of degree $k_n$ is given by $\widetilde{T}_{\text{box}}(k_n, C)$. Therefore, in order to prove Proposition \ref{prop:concentration_local_clustering_P_n} it suffices to show that $\widetilde{T}_{\text{box}}(k_n,C)$ is sufficiently concentrated around its mean. This is done in the following proposition.

%\subsection{The proof of Proposition \ref{prop:concentration_local_clustering_P_n}}
%
%We start with the concentration result for $\widetilde{T}_{\text{box}}(k_n)$.

\begin{proposition}[Concentration $\widetilde{T}_{\text{box}}(k_n,C)$]\label{prop:concentration_tilde_T_P_n}
Let $\alpha > \frac{1}{2}$, $\nu > 0$ and let $(k_n)_{n \ge 1}$ be any positive sequence satisfying $k_n = \smallO{n^{\frac{1}{2\alpha+1}}}$. Then for any $C > 0$, as $n \to \infty$,
\[
	\Exp{\widetilde{T}_{\text{box}}(k_n,C)^2} = \left(1 + \smallO{1}\right)\Exp{\widetilde{T}_{\text{box}}(k_n, C)}^2.
\]
\end{proposition}

We postpone the proof of this proposition till Section~\ref{ssec:concentration_tilde_T} and first use it to prove Proposition~\ref{prop:concentration_local_clustering_P_n}.

\begin{proof}[Proof of Proposition \ref{prop:concentration_local_clustering_P_n}]
We bound the expectation as follows,
\begin{align*}
	\Exp{\left|c^\ast(k_n; \Gbox) - \Exp{c^\ast(k_n; \Gbox)}\right|} 
	&\le \frac{\Exp{\left|\widetilde{T}_{\text{box}}(k_n,C) - \Exp{\widetilde{T}_{\text{box}}(k_n,C)}\right|}}
		{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}}\\
	&\hspace{10pt} + 2 \Exp{\left|c^\ast(k_n; \Gbox) - \frac{\widetilde{T}_{\text{box}}(k_n,C)}{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}}\right|}.
\end{align*}
We will show that both terms are $\smallO{s(k_n)}$.

First we note that $\ind{p_2 \in \BallPo{p_1} \cap \Rcal} \le \ind{p_2 \in \BallPon{p_1}}$ and hence $\widetilde{T}_{\text{box}}(p) \le T_{\text{box}}(p)$. This implies that
\[
	 \frac{\widetilde{T}_{\text{box}}(k_n,C)}{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}} \le c^\ast(k_n; \Gbox). 
\]
and therefore
\[
	\Exp{\left|c^\ast(k_n; \Gbox) - \frac{\widetilde{T}_{\text{box}}(k_n,C)}{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}}\right|}
	= \Exp{c^\ast(k_n; \Gbox)} 
	- \frac{\Exp{\widetilde{T}_{\text{box}}(k_n,C)}}{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}}.
\]
For the expectation of $\widetilde{T}_{\text{box}}(k_n,C)$ we use that 
\[
	\CExp{\widetilde{T}_{\text{box}}(p)}{\Dbox(p) = k_n}
= \binom{k_n}{2} \Mu{\BallPon{y}}^{-2} \Exp{\widetilde{T}_{\text{box}}(p)},
\] to get
\begin{align*}
	\Exp{\widetilde{T}_{\text{box}}(k_n,C)} 
	&= \int_{\Kcal_{C,n}(k_n)} \CExp{\widetilde{T}_{\text{box}}(p)}{\Dbox(p) = k_n}
		\rho_{\text{box}}(y,k_n) f(x,y) \dd x \dd y\\
	&= (1+\smallO{1}) \binom{k_n}{2} \int_{\Kcal_{C,n}(k_n)} \Mu{\BallPon{y}}^{-2} \Exp{\widetilde{T}_{\text{box}}(y)}
		\rho_{\text{box}}(y,k_n) \alpha e^{-\alpha y} \dd y\\
	&= (1+\smallO{1}) \frac{1}{2} \int_{\Kcal_{C,n}(k_n)} \Exp{\widetilde{T}_{\text{box}}(y)}
			\rho_{\text{box}}(y,k_n) \alpha e^{-\alpha y} \dd y\\
	&= (1+\smallO{1}) n \binom{k_n}{2} \int_0^\infty P(y) \alpha e^{-\alpha y} \dd y,
\end{align*}
where the last line is due to Corollary~\ref{cor:adjusted_triangle_counting_P_n}. Since $\Exp{N_{\text{box}}(k_n)} = (1+\smallO{1}n p_{k_n}$ it follows that
\[
	\frac{\Exp{\widetilde{T}_{\text{box}}(k_n,C)}}{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}}
	= (1+\smallO{1}) \frac{\int_0^\infty P(y) \alpha e^{-\alpha y} \dd y}{p_{k_n}}
	= (1 + \smallO{1}) \gamma(k_n).
\]
On the other hand, Proposition~\ref{prop:convergence_average_clustering_P_n} implies that $\Exp{c^\ast(k_n; \Gbox)} = (1+\smallO{1})\gamma(k_n)$ and thus we conclude that
\[
	2\Exp{\left|c^\ast(k_n; \Gbox) - \frac{\widetilde{T}_{\text{box}}(k_n,C)}{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}}\right|}
	= \smallO{s(k_n)}.	
\]

For the remaining term we use Proposition \ref{prop:concentration_tilde_T_P_n} to obtain
\begin{align*}
	\Exp{\left|\widetilde{T}_{\text{box}}(k_n,C) - \Exp{\widetilde{T}_{\text{box}}(k_n,C)}\right|}
	&\le \left(\Exp{\widetilde{T}_{\text{box}}(k_n,C)^2} 
		- \Exp{\widetilde{T}_{\text{box}}(k_n,C)}^2\right)^{\frac{1}{2}}\\
	&= \smallO{\Exp{\widetilde{T}_{\text{box}}(k_n,C)}}.
\end{align*}
This implies
\begin{align*}
	\frac{\Exp{\left|\widetilde{T}_{\text{box}}(k_n,C) - \Exp{\widetilde{T}_{\text{box}}(k_n,C)}\right|}}
		{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}}
	&= \smallO{\frac{\Exp{\widetilde{T}_{\text{box}}(k_n,C)}}{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}}}
	= \smallO{s(k_n)},
\end{align*}
which finishes the proof.
\end{proof}

@@

Recall (see \eqref{eq:def_degree_triangle_count_in_K}) that for any $C > 0$
\[
	\widetilde{T}_{\text{box}}(k,C) = \sum_{p \in \Pcal_n \cap \Kcal_{C,n}(k_n)} \ind{\text{deg}_{\text{box}}(p) = k} \widetilde{T}_{\text{box}}(p)
\]
Then we have
\[
	\widetilde{T}_{\text{box}}(k,C)^2 = \hspace{-5pt} \sum_{p, p^\prime \in \Pcal_n \cap \Kcal_C(k)}
		\hspace{-3pt} \ind{D_{\Pcal,n}(p), \, D_{\Pcal,n}(p^\prime) = k} 
		\sum_{(p_1, p_2), (p_1^\prime, p_2^\prime) \in \Pcal_n \setminus (0,y)}^{\ne} \hspace{-3pt}
		\widetilde{T}_{\Pcal}(p,p_1,p_2) \widetilde{T}_{\Pcal}(p^\prime, p_1^\prime, p_2^\prime),
\]
This expression can be written as the sums of several terms, depending on how $\{p, p_1, p_2\}$ and $\{p^\prime, p_1^\prime, p_2^\prime\}$ intersect. To this end we define, for $a \in \{0,1\}$ and $b \in \{0,1,2\}$,
\[
	I_{a,b} = \hspace{-3pt} \sum_{p, p^\prime \in \Pcal_n \cap \Kcal_C(k) \atop |\{p\} \cap \{p^\prime\}| = a}
	\hspace{-5pt} \ind{D_{\Pcal,n}(p), \, D_{\Pcal,n}(p^\prime) = k} J_b(p,p^\prime),
\]
where
\[
	J_b(p,p^\prime) = \hspace{-10pt} \sum_{p_1, p_2, p_1^\prime, p_2^\prime \in \Pcal_n 
		\atop |\{p_1, p_2\} \cap \{p_1^\prime, p_2^\prime\}| = b}^{\ne}
		\hspace{-5pt} T_{\Pcal,n}(p,p_1,p_2) T_{\Pcal,n}(p^\prime, p_1^\prime, p_2^\prime),
\]
with the summation being over all two distinct pairs $(p_1, p_2)$ and $(p_1^\prime, p_2^\prime)$. Then we have
\[
	\widetilde{T}_{\text{box}}(k, C)^2 = \sum_{a = 0}^1 \sum_{b = 0}^2 I_{a,b}.
\]

To prove Proposition \ref{prop:concentration_tilde_T_P_n} we will deal with each of the $I_{a,b}$ separately, showing that 
\begin{equation}\label{eq:variance_T_I_00}
	\Exp{I_{0,0}} = (1+\smallO{1})\Exp{\widetilde{T}_{\text{box}}(k_n)}^2
\end{equation}
and for all other combinations
\begin{equation}\label{eq:variance_T_I_ab}
	\Exp{I_{a,b}} = \smallO{\Exp{\widetilde{T}_{\text{box}}(k_n)}^2}.
\end{equation}
Note that $J_{b}(p,p^\prime) \le J_{0}(p,p^\prime)$ and, since $I_{1,2} = \widetilde{T}_{\text{box}}(k_n,C)$, \eqref{eq:variance_T_I_ab} holds for $I_{1,2}$.

@@

Throughout this proof we set $i = |\{p^\prime, p_1, p_2, p_1^\prime, p_2^\prime\} \cap \BallPon{p}|$, $j = |\{p^\prime\} \cap \BallPon{p}|$ and define $i^\prime, j^\prime$ in a similar way by interchanging the primed and non-primed variables. In addition, we write $D_{\Pcal,n}(p,p^\prime,k,\ell)$ to denote the indicator that $|\BallPon{p} \cap (\Pcal_n \setminus \{p, p^\prime, p_1, p_2, p_1^\prime, p_2^\prime\})| = k$ and $|\BallPon{p^\prime} \cap (\Pcal_n \setminus \{p, p^\prime, p_1, p_2, p_1^\prime, p_2^\prime\})| = \ell$. Then, 
\begin{align*}
	&\Exp{\ind{D_{\Pcal,n}(p) = k_n, D_{\Pcal,n}(p^\prime) = k_n} J_b(p,p^\prime)}\\
	&= \Exp{\sum_{p_1, p_2, p_1^\prime, p_2^\prime \in \Pcal_n 
		\atop |\{p_1, p_2\} \cap \{p_1^\prime, p_2^\prime\}| = b}^{\ne}
			\hspace{-7pt} D_{\Pcal,n}(p,p^\prime, k_n-i,k_n- i^\prime) \,
			\widetilde{T}_{\text{box}}(p,p_1,p_2) \widetilde{T}_{\text{box}}(p^\prime, p_1^\prime, p_2^\prime)},
\end{align*}
where the sum is over all distinct pairs $(p_1, p_2)$ and $(p_1^\prime, p_2^\prime)$. We also that 
\[
	\Exp{T_\Pcal(k_n)} = \bigT{n k_n^{-(2\alpha - 1)} s_\alpha(k_n)}.
\]

@@

%	&\le \int_{\mathcal{E}_{C}(k_n)^c} \rho_{n}(y,k_n) \Exp{\widetilde{T}_{\text{box}}(p)}
%		\Exp{\widetilde{T}_{\text{box}}(p^\prime)} 

@@

%	&\le \bigO{\log(n)^2} \left(\int_{\Kcal_{C}(k_n)} \rho_{n}(y,k_n) \Exp{\widetilde{T}_{\text{box}}(p)} 

@@

%	\mathcal{J}_1^\ast = \int_{\Rcal^2} \ind{0\le y_1, y_2 \le h(p,p^\prime)} \widetilde{T}_{\text{box}}(p,p_1,p_2) 

@@

%	\mathcal{J}_1^\ast = (1+\smallO{1})\Exp{\widetilde{T}_{\text{box}}(p)} = (1 + \smallO{1})\binom{k_n}{2} \Delta_\Pcal(y),

@@

By Lemma~\ref{lem:joint_degree_distribution_shift} 
\[
	\rho_(p,p^\prime, k_n-i,k_n-i^\prime) = (1+o(1))\rho(p,p^\prime, k_n-j,k_n-j^\prime) 
	= (1+o(1))\rho_{n}(p, p^\prime, k_n, k_n)
\]
which now no longer depends on the other four points $p_1, p_2, p_1^\prime, p_2^\prime$. Hence, using the Campbell-Mecke formula, we get
\[
	\Exp{I_{0,0}^\ast} = (1 + \smallO{1})\int_{\mathcal{E}_{\varepsilon}(k_n)}\rho_{n}(p,p^\prime,k_n, k_n)
		\Exp{\widetilde{T}_{\text{box}}(p)}\Exp{\widetilde{T}_{\text{box}}(p^\prime)} f(x,y)
		f(x^\prime,y^\prime) \, dx^\prime \, dx \, dy^\prime \, dy,
\]
